---
title: 'Reproducible Research: Peer Assessment 1'
author: "Marinus Koekemoer"
date: "20 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

1. Load the data

```{r}
Act_Mon_Data <- read.csv("activity.csv")
```

2. Process/transform the data

```{r}
StepsPerDay <- aggregate(steps ~ date, data=Act_Mon_Data, FUN=sum)
```

## What is mean total number of steps taken per day?

1. Histogram of the total number of steps taken each day

```{r}
hist(StepsPerDay$steps)
```

2. Mean and median number of steps taken each day

```{r}
mean(StepsPerDay$steps)
median(StepsPerDay$steps)
```

## What is the average daily activity pattern?

1. Time series plot of the average number of steps taken

```{r}
StepsPerInterval<-aggregate(steps~interval,data=Act_Mon_Data,mean,na.rm=TRUE)
plot(steps~interval,data=StepsPerInterval,type="l")
```

2. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
StepsPerInterval$interval[which.max(StepsPerInterval$steps)]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(Act_Mon_Data$steps))
```

2. Strategy for filling in all of the missing values in the dataset.

Overall Mean stpes will be used

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
ImputedAct_Mon_Data <-Act_Mon_Data
ImputedAct_Mon_Data$steps[is.na(ImputedAct_Mon_Data$steps)] <- mean(ImputedAct_Mon_Data$steps, na.rm=TRUE)
```

4. Histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
ImputeStepsPerDay <- aggregate(steps ~ date, data=ImputedAct_Mon_Data, FUN=sum)
hist(ImputeStepsPerDay$steps)
```

```{r}
mean(ImputeStepsPerDay$steps)
median(ImputeStepsPerDay$steps)
```

Diference in the means are
```{r}
mean(ImputeStepsPerDay$steps) - mean(StepsPerDay$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
ImputedAct_Mon_Data$day=ifelse(as.POSIXlt(as.Date(ImputedAct_Mon_Data$date))$wday%%6==0,"weekend","weekday")
ImputedAct_Mon_Data$day=factor(ImputedAct_Mon_Data$day,levels=c("weekday","weekend"))
```

2. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
stepsWeek=aggregate(steps~interval+day,ImputedAct_Mon_Data,mean)
library(lattice)
xyplot(steps~interval|factor(day),data=stepsWeek,aspect=1/2,type="l")
```
